Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-10 manuscript.Rmd
recording stats

```{r participants, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
ppts <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")))
num_younger <- ppts %>% filter(group == "Younger") %>% .$id %>% unique(.) %>% length(.)
num_older <-  ppts %>% filter(group == "Older") %>% .$id %>% unique(.) %>% length(.)
num_sessions <- ppts %>% group_by(group, id, session) %>% slice_head(n = 1) %>% count %>% group_by(group, id) %>% count
age <- ppts %>% group_by(group) %>% get_summary_stats(age) %>% column_to_rownames("group")
```

@@

end_times <- ppts %>% group_by(id_uni) %>% slice_max(clock_time_start) %>% pull(clock_time_start) %>% sort
end_times <- str_sub(as.character(end_times), start = 1, end = -4)

coverage <- ppts %>% mutate(missing = ifelse(is.na(nap_time), 1, 0)) %>%
  group_by(id_uni) %>% 
  summarize(blocks = n(), missing = sum(missing, na.rm = T), naps = sum(nap_time, na.rm = T)) %>% 
  mutate(total = blocks * 5/60, prop_accounted = 1-(missing/blocks), naptime = naps * 5)
```

After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house. The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, making us confident that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.

This leads to our last questions---how successfully did recordings capture infants' entire day, and how suitable were daylong recordings for capturing individual differences in body position? Figure \@ref(fig:timelines) depicts body position timelines across the day for each session, divided into younger and older infants, predicted from the group model. The start time ranged from `r start_times[1]` to `r start_times[length(start_times)]` with a median of `r start_times[floor(length(start_times)/2)]`. Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended.

>>>>

```{r participants, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
ppts <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older"))) %>% 
  filter(id_uni != 10401)
num_younger <- ppts %>% filter(group == "Younger") %>% .$id %>% unique(.) %>% length(.)
num_older <-  ppts %>% filter(group == "Older") %>% .$id %>% unique(.) %>% length(.)
num_sessions <- ppts %>% group_by(group, id, session) %>% slice_head(n = 1) %>% count %>% group_by(group, id) %>% count
age <- ppts %>% group_by(group) %>% get_summary_stats(age) %>% column_to_rownames("group")
```

@@

end_times <- ppts %>% filter(id_uni != 11104, id_uni != 801) %>%
  group_by(id_uni) %>% slice_max(clock_time_start) %>% pull(clock_time_start) %>% sort
end_times <- str_sub(as.character(end_times), start = 1, end = -4)

coverage <- ppts %>% mutate(missing = ifelse(is.na(nap_time), 1, 0)) %>%
  filter(id_uni != 11104, id_uni != 801) %>% 
  group_by(id_uni) %>% 
  summarize(blocks = n(), missing = sum(missing, na.rm = T), naps = sum(nap_time, na.rm = T)) %>% 
  mutate(total = blocks * 5/60, 
         prop_accounted = 1-(missing/(blocks-missing)), 
         hours_recorded = (blocks-missing-naps)*5/60,
         naptime = naps * 5/60)
```

After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house. The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.

This leads to two final questions---how successfully did recordings capture infants' entire day, and how suitable were day-long recordings for capturing individual differences in body position? Figure \@ref(fig:timelines) depicts body position timelines across the day for each session, divided into younger (A) and older infants (B). Predictions from the group model were used to ensure that motion data were classified consistently across all sessions. Session start times ranged from `r start_times[1]` to `r start_times[length(start_times)]` with a median of `r start_times[floor(length(start_times)/2)]`. Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group). With two exceptions, recordings lasted until the infants' bed time. Older participant \#1 had the equipment picked up on the same day rather than the next day, so the recording ended at 17:00. Older participant \#11 wore the equipment in the morning, but the family left the house at 10:15 and remained out for the rest of the day, choosing not to put the equipment back on when they returned home in the late afternoon. Among those participants who wore the equipment until bedtime, recordings ended between `r end_times[1]` and `r end_times[length(end_times)]` with a median of `r end_times[floor(length(end_times)/2)]`.

Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)). The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours. Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours. Of the remaining awake time during the recording period, body position data was collecting during proportions that ranged from `r apa_num(min(coverage$prop_accounted))`-`r apa_num(max(coverage$prop_accounted))` hours with a median of `r apa_num(median(coverage$prop_accounted))`. Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home. Overall, the recordings produced a median of `r apa_num(median(coverage$hours_recorded))` hours of motion data, with the entire dataset totaling `r apa_num(sum(coverage$hours_recorded))` hours.

### 2023-03-10 manuscript.tex
recording stats

Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{Older} infants were between 11 and 14 months. There were 9 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female). Families were recruited through social media advertisements and from community events in Southern California. The ethnicity for infants was reported as X Hispanix/Latinx and X for not Hispanix/Latinx. Race was reported as X Caucasian, X Asian, X \textasciitilde\textasciitilde\textasciitilde{} and X others. Parents were compensated \$30 in cash for each visit they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All participants gave their informed consent to participate.

Most participants were tested in a single session (\emph{n} = 17), but 6 participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.84 to 7.16 (\emph{M} = 4.99) and older infants' age ranged from 10.74 to 14.23 (\emph{M} = 11.75).

@@

We report four sets of results based on 34 full-day testing sessions resulting in a total of 301.74 hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated on the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is crucial, because it provides the first ever test of how well body classification models predict behavior beyond the initial, supervised period. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day should be valid. In the third set of results we examine the data quality of full-day recordings, and in the fourth set of results we test whether full-day recordings capture well-established age differences in body position experiences.

@@

Whereas the first set of analyses used all 34 sessions, this was not possible in the distal comparison. Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period. Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording. Three additional sessions were excluded because a caregiver accidentally turned off the video camera (\emph{n} = 1) or left the house (\emph{n} = 2). This left 22 sessions with usable distal comparison data.

@@

After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house. The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, making us confident that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.

This leads to our last questions---how successfully did recordings capture infants' entire day, and how suitable were daylong recordings for capturing individual differences in body position? Figure \ref{fig:timelines} depicts body position timelines across the day for each session, divided into younger and older infants, predicted from the group model. The start time ranged from 08:55 to 13:20 with a median of 10:20. Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended.

>>>>

Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{Older} infants were between 11 and 14 months. There were 8 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female). Families were recruited through social media advertisements and from community events in Southern California. The ethnicity for infants was reported as X Hispanix/Latinx and X for not Hispanix/Latinx. Race was reported as X Caucasian, X Asian, X \textasciitilde\textasciitilde\textasciitilde{} and X others. Parents were compensated \$30 in cash for each visit they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All participants gave their informed consent to participate.

Most participants were tested in a single session (\emph{n} = 16), but 6 participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 33 sessions, with 13 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.84 to 7.16 (\emph{M} = 5.03) and older infants' age ranged from 10.74 to 14.23 (\emph{M} = 11.75).

@@

We report four sets of results based on 33 full-day testing sessions resulting in a total of 294.07 hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated on the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is crucial, because it provides the first ever test of how well body classification models predict behavior beyond the initial, supervised period. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day should be valid. In the third set of results we examine the data quality of full-day recordings, and in the fourth set of results we test whether full-day recordings capture well-established age differences in body position experiences.

@@

Whereas the first set of analyses used all 33 sessions, this was not possible in the distal comparison. Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period. Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording. Three additional sessions were excluded because a caregiver accidentally turned off the video camera (\emph{n} = 1) or left the house (\emph{n} = 2). This left 22 sessions with usable distal comparison data.

@@

After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house. The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.

This leads to two final questions---how successfully did recordings capture infants' entire day, and how suitable were day-long recordings for capturing individual differences in body position? Figure \ref{fig:timelines} depicts body position timelines across the day for each session, divided into younger (A) and older infants (B). Predictions from the group model were used to ensure that motion data were classified consistently across all sessions. Session start times ranged from 08:55 to 13:20 with a median of 10:20. Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group). With two exceptions, recordings lasted until the infants' bed time. Older participant \#1 had the equipment picked up on the same day rather than the next day, so the recording ended at 17:00. Older participant \#11 wore the equipment in the morning, but the family left the house at 10:15 and remained out for the rest of the day, choosing not to put the equipment back on when they returned home in the late afternoon. Among those participants who wore the equipment until bedtime, recordings ended between 17:25 and 22:15 with a median of 19:20.

Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}). The total length of the recording period ranged from 6.92-11.75 hours with a median of 9.08 hours. Nap times reported during the recording period ranged from 0.00-4.99 hours with a median of 2.22 hours. Of the remaining awake time during the recording period, body position data was collecting during proportions that ranged from 0.18-1.00 hours with a median of 1.00. Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home. Overall, the recordings produced a median of 6.15 hours of motion data, with the entire dataset totaling 201.06 hours.
